Elicitation Complexity via Convex Piecewise Linear Loss Functions

### 2018-10-15 jessie-writeup/jessie-writeup.tex
Cyclic proof of TFAE and finished up a few thoughts

A set $\Theta \subseteq 2^{\Delta(\Y)}$ is an unlabeled property, and finite if $|\Theta| < \infty$.
Observe that $\Theta$ is a set of level sets, and not a function.\raft{Careful: $\Theta$ does not need to divide anything; $\Theta = \{\emptyset\}$ is technically an unlabeled (degenerate) property}

@@

\begin{notation}
  Consider two labeled properties $\Gamma' : \P \toto \R$ and $\Gamma : \P \toto \R$.
  We say $\Gamma' \preceq \Gamma$ if removing reports from $\Gamma$ yields $\Gamma'$, that is, if $\Gamma' = \Gamma \cap \R'$ for some $\R' \subseteq \R$.
  \jessiet{Changed $\R \subseteq \R$ to $\R' \subseteq \varphi(\R)$.}
\end{notation}

@@

\begin{definition}\label{def:unlabeled-nonredundant}
  An unlabeled property $\Theta$ is \emph{nonredundant} if for there are no full dimensional level sets $\theta_1, \theta_2 \in \Theta$ so that $\theta_1 \subseteq \theta_2$.
  \raft{Have not defined elicitability for unlabeled properties.  Also, what does ``almost everywhere'' mean in this context, since $\Theta$ is just a set of sets?  I think we just want to say that there are no distinct $\theta_1,\theta_2 \in \Theta$ such that $\theta_1 \subseteq \theta_2$.- you're right... shoot}
  \jessie{Should be equivalent to definition of nonredundancy for labeled property $\Gamma$, shown in Lemma~\ref{lem:nonredundant-defs-equivalent}.}
\end{definition}

Other notes before starting:
\begin{itemize}
\item (Assuming the metric space is $(\Delta(\Y), \rho)$, where $\rho$ is Euclidean distance, although I don't think the choice of metric matters?).
We just use this looking at interiors of level sets.
\item Using generalized Lebesgue measure on $\P \subset \reals^{|\Delta(\Y)|-1}$
\item The interior of a level set $X_r$ is defined as $\inter{X_r} = \{p \in X_r: \epsilon > 0,  \forall q \in B(\epsilon,p), r \in \argmin_r \E_q L(r,Y)\}$
\item Define the boundary of the closed set $X$ by $bd(X) := X \setminus \inter{X}$.
\item A level set $\theta$ is \emph{full-dimensional} if it has nonempty interior. %% old: $dim(\aff(\theta)) = dim(\P)$. 
%Equivalently, $\theta$ is full-dimensional if there are distributions $p_1, p_2, \ldots, p_n \in \theta$ so that $p_1, \ldots, p_n$ are affinely independent.
\raft{Careful; I think you need convexity here.  Probably you want to define full-dimensional as having nonempty interior, and prove this as a consequence of convexity.}
\jessiet{What is the formal, ``official'' definition of full-dimensional, if one exists?}
%\item A full-dimensional level set has non-empty interior.
\end{itemize}

@@

  The following are equivalent:
  \begin{enumerate}
  \item There is a finite set $\Theta \subseteq \strip(\Gamma)$ of full-dimensional level sets, such that: (a) $\bigcup_{\theta\in\Theta} \theta = \P$, (b) for all $\theta \in \Theta$, there exists an $r_\theta \in \R$ such that $\Gamma(p) = \{r_\theta\}$ for all $p \in \inter{\theta}$.
  \item $\trim(\Gamma)$ is finite.
  \item $\strip(\Gamma)$ is finite.
  \item \jessie{Added after fplc-elic.tex}
	\begin{itemize}
		\item[4.] $\trim(\Gamma)$ is finite and elicitable.
		\item[5.] There is a finite set of full dimensional level sets of $\Gamma$ that union to $\P$.
		\item[6.] There is a nondegenerate, finite property $\Gamma':\P\toto\R'$ such that there is an injection $\varphi:\R'\to\R$ such that $r\in\gamma(p) \iff \varphi(r) \in \Gamma(p)$. (i.e. $\Gamma$ embeds finite $\Gamma'$)
	\end{itemize}
  \end{enumerate}
\end{proposition}

\begin{proof}
First, we show $2 \iff 3$ then proceed in a cycle to show $6 \implies 1 \implies 5 \implies 4 \implies 2 \implies 6$.

@@

	\item[$2 \iff 3$] 
	We know $\trim(\Gamma) \subseteq \strip(\Gamma)$, and if $\strip(\Gamma)$ is finite, then any subset of $\strip(\Gamma)$ is finite.
	Thus we just show $2 \implies 3$.

@@

	Since $\Gamma$ is elicitable, the score $G(p)$ is concave, thus differentiable almost everywhere (in the simplex).
	Thus, every level set with nonempty interior has $|\partial G(p)| = |\Gamma(p)| = 1$ for all $p \in \inter{\theta}$, and there is then a report $r \in \R'$ so that $\theta \in \Gamma_r$.

@@

\item[$6 \implies 1$] 
Let the elicitable property $\Gamma$ embed some finite property $\gamma$.
Then since $\gamma$ is a finite set of level sets, and every level set of $\Gamma$ is a subset of a level set of $\gamma$ by Conjecture~\ref{conj:lev-sets-subsets}.

@@

Therefore the set $\Theta = \{\gamma_r : r \in \R' \} = \trim(\Gamma) \subseteq \strip(\Gamma)$ is a finite set of full dimensional level sets that union to $\P$ with a unique optimal report on the interior of each level set. 
\item[$1 \implies 5$] Trivial
\item[$5 \implies 4$]
Let $\Theta$ be a set of full-dimensional level sets of $\Gamma$ that union to $\P$.
First, we know $\Gamma$ is nondegenerate.

@@

Therefore $\trim(\Gamma)$ is finite, and $L|_\R$ elicits $\trim(\Gamma)$ from \emph{cvx-fin-notes.tex}.

\item[$4 \implies 2$] Trivial
\item[$2 \implies 6$] Let $\trim(\Gamma)$ be finite.
Then there is a nondegenerate, nonredundant, finite property $\gamma$ so that $\{ \gamma_r : r \in \R' \} = \strip(\gamma) = \trim(\Gamma)$.

Then we want to show $\Gamma$ embeds $\gamma$.
Since we know $\card(\trim(\Gamma)) = \card(\R')$ is finite and $\trim(\Gamma)$ is a subset of the codomain of $\Gamma$, $\R$, then $\card(\R')) \leq \card(\R) \jessie{?= \card(\reals^d)}$, there is an injection $\varphi : \R' \to \R$ from Folland Section 0.

Consider some $r \in \gamma(p)$.
Since there is an injection $\varphi$ from $\R' \to \R$, then $r \in \gamma(p) \implies \varphi(r) \in \varphi(\gamma(p)) = \Gamma(p)$.
\jessie{Hand wavvy there}

@@

Therefore, $2 \implies 6$.

@@

\begin{lemma}\label{lem:pos-measure-iff-nonempty-int}
	Let $S$ be a convex set.
	$S$ has nonempty interior $\iff S$ has positive Lebesgue measure.
\end{lemma}
\begin{proof}

	\begin{enumerate}
		\item[$\implies$]
		Let $S$ have nonempty interior.
		We know that as $\inter{S}$ and $bd(S)$ are disjoint, $\mu(S) = \mu(\inter{S}) + \mu(bd(S))$.
		Since the interior of $S$ is nonempty, there is an open ball $B \subseteq S$ such that $x \in B \implies x \in \inter S$.
		As Lebesgue measure in $\reals^n$ is a product measure and every open interval has positive measure, then the Lebesgue measure of the ball $B$ is the product of finitely many positive measures, and therefore positive.
		As $0 < \mu(B) \leq \mu(\inter S) \leq \mu(S)$ by monotonicity, we see that $S$ with nonempty interior has positive measure.

		\item[$\impliedby$]  Follows from the contrapositive and definition of measure.
		That is, $\mu(S) = \mu(\inter{S}) + \mu(bd(S)) = \mu(\inter{S})$.\footnote{Since $S$ is convex, $\mu(bd(S)) = 0$ according to \url{https://math.stackexchange.com/questions/613408/when-does-the-boundary-have-measure-zero}.}
		As $\inter{S} = \emptyset \implies \mu(\inter{S}) = 0$, and hence does not have positive measure.
		\jessie{This isn't right-- I definitely messed up the contrapositive here.  This was where we talked about showing $\theta$ is a convex level set with nonempty interior iff there are $x_1, \ldots, x_n \in \theta$ are affinely independent.}
	\end{enumerate}
\end{proof}

\begin{lemma}
	Let $\theta$ be a [closed] convex level set.
	$\theta$ is full dimensional $\iff$ there are distributions $p_1, p_2, \ldots, p_n \in \theta$ so that $p_1, p_2, \ldots, p_n$ are affinely independent.
\end{lemma}
\begin{proof}
	Recall that $n := |\Delta(\Y)|$, so we say the simplex is $n$-dimensional.
	\begin{enumerate}
		\item [$\implies$]
		Assume the level set $\theta$ is full dimensional.
		Then there exists a $p \in \inter{\theta}$ and for every $q \in B(\epsilon, p)$, we know that $q \in \theta$.
		Take $q_i = p + \epsilon e_i$\footnote{$e_i$ is the vector with $1$ in the $i^{th}$ component, and is $0$ elsewhere} in each of the $n$ dimensions, and since $\theta$ is full dimensional and closed, we know each $q_i \in\theta$.
		As $\{e_i\}_{i=1}^n$ is affinely independent (it forms the standard basis for $\reals^n$), so then is $\{q_i\}_{i=1}^n$ as we are invariant to additive and multiplicative shifts.

@@

\begin{conjecture}\label{conj:prec-if-embeds}
	If the elicitable property $\Gamma$ embeds the nondegenerate finite $\gamma$, then $\gamma \preceq \Gamma$ and $\trim(\Gamma) = \strip(\gamma)$.
\end{conjecture}
\begin{proof}

@@

	To show $\gamma \preceq \Gamma$, we know that since $\Gamma$ embeds $\gamma$, there is an injection $\varphi : \R' \to \R$ so that $r' \in \gamma(p) \iff \varphi(r') \in \Gamma(p)$.
	As $\varphi(\R') \subseteq \R$ and $\gamma$ is nondegenerate, we know that for all $r \in \R'$

	Therefore, $\gamma(p) = \Gamma(p) \cap \R'$

	Finally, we show $\gamma$ is nonredundant. \jessie{TODO...}

>>>>

A set $\Theta \subseteq 2^{\Delta(\Y)}$ is an unlabeled property, and finite if $|\Theta| < \infty$.
Observe that $\Theta$ is a set of level sets, and not a function.
\raft{Careful: $\Theta$ does not need to divide anything; $\Theta = \{\emptyset\}$ is technically an unlabeled (degenerate) property JF: Changed}

@@

\begin{notation}
  Consider two labeled properties $\Gamma' : \P \toto \R$ and $\Gamma : \P \toto \R$.
  We say $\Gamma' \preceq \Gamma$ if removing reports from $\Gamma$ yields $\Gamma'$, that is, if $\Gamma' = \Gamma \cap \varphi(\R')$ for some $\varphi(\R') \subseteq \R$.
  \jessiet{Changed $\R' \subseteq \R$ to $\varphi(\R') \subseteq \R$.}
\end{notation}

@@

\begin{definition}\label{def:unlabeled-nonredundant}
  An unlabeled property $\Theta$ is \emph{nonredundant} if for there are no full dimensional level sets $\theta_1, \theta_2 \in \Theta$ so that $\theta_1 \subseteq \theta_2$.
  \raft{Have not defined elicitability for unlabeled properties.  Also, what does ``almost everywhere'' mean in this context, since $\Theta$ is just a set of sets?  I think we just want to say that there are no distinct $\theta_1,\theta_2 \in \Theta$ such that $\theta_1 \subseteq \theta_2$.  JF: you're right... shoot}
  \jessie{Should be equivalent to definition of nonredundancy for labeled property $\Gamma$, shown in Lemma~\ref{lem:nonredundant-defs-equivalent}.}
\end{definition}

Other notes before starting:
\begin{itemize}
\item (Assuming the metric space is $(\Delta(\Y), \rho)$, where $\rho$ is Euclidean distance, although I don't think the choice of metric matters?).
We just use this looking at interiors of level sets.
\item Using generalized Lebesgue measure on $\P \subset \reals^{|\Delta(\Y)|-1}$
\item The interior of a level set $X_r$ is defined as $\inter{X_r} = \{p \in X_r: \epsilon > 0,  \forall q \in B(\epsilon,p), r \in \argmin_r \E_q L(r,Y)\}$
\item Define the boundary of the closed set $X$ by $bd(X) := X \setminus \inter{X}$.
\item A convex level set $\theta$ is \emph{full-dimensional} if it has nonempty interior. %% old: $dim(\aff(\theta)) = dim(\P)$. 
%Equivalently, $\theta$ is full-dimensional if there are distributions $p_1, p_2, \ldots, p_n \in \theta$ so that $p_1, \ldots, p_n$ are affinely independent.
\raft{Careful; I think you need convexity here.  Probably you want to define full-dimensional as having nonempty interior, and prove this as a consequence of convexity.}
\jessiet{What is the formal, ``universal'' definition of full-dimensional, if one exists?}
%\item A full-dimensional level set has non-empty interior.
\end{itemize}

@@

  The following are equivalent:
  \begin{enumerate}
  \item There is a nondegenerate, finite property $\Gamma':\P\toto\R'$ such that there is an injection $\varphi:\R'\to\R$ such that $r\in\gamma(p) \iff \varphi(r) \in \Gamma(p)$. (i.e. $\Gamma$ embeds finite $\Gamma'$)
  \item There is a finite set of full dimensional level sets of $\Gamma$ that union to $\P$.
  \item There is a finite set $\Theta \subseteq \strip(\Gamma)$ of full-dimensional level sets, such that: (a) $\bigcup_{\theta\in\Theta} \theta = \P$, (b) for all $\theta \in \Theta$, there exists an $r_\theta \in \R$ such that $\Gamma(p) = \{r_\theta\}$ for all $p \in \inter{\theta}$.
  \item $\strip(\Gamma)$ is finite.
  \item $\trim(\Gamma)$ is finite.
  \item $\trim(\Gamma)$ is finite and elicitable.

  \end{enumerate}
\end{proposition}

\begin{proof}
First, we show $5 \iff 4$ then proceed in a cycle to show $1 \implies 3 \implies 2 \implies 6 \implies 5 \implies 1$.

@@

	\item[$5 \iff 4$] 
	We know $\trim(\Gamma) \subseteq \strip(\Gamma)$, and if $\strip(\Gamma)$ is finite, then any subset of $\strip(\Gamma)$ is finite.
	Thus we just show $5 \implies 4$.

@@

	Since $\Gamma$ is elicitable, the score $G(p)$ is concave, thus differentiable almost everywhere in the probability space $\P$.
	Thus, every level set with nonempty interior has $|\partial G(p)| = |\Gamma(p)| = 1$ for all $p \in \inter{\theta}$, and there is then a report $r \in \R'$ so that $\theta \in \Gamma_r$.

@@

\item[$1 \implies 3$] 
Let the elicitable property $\Gamma$ embed some finite property $\gamma$.
Then since $\gamma$ is a finite set of level sets, and every level set of $\Gamma$ is a subset of a level set of $\gamma$ by Conjecture~\ref{conj:lev-sets-subsets}.

@@

Therefore the set $\Theta = \{\gamma_r : r \in \R' \} = \trim(\Gamma) \subseteq \strip(\Gamma)$ is a finite set of full dimensional level sets that union to $\P$ with a unique optimal report on the interior of each level set. 
\item[$3 \implies 2$] Trivial
\item[$2 \implies 6$]
Let $\Theta$ be a set of full-dimensional level sets of $\Gamma$ that union to $\P$.
First, we know $\Gamma$ is nondegenerate.

@@

Therefore $\trim(\Gamma)$ is finite, and $L|_{\R'}$ elicits $\trim(\Gamma)$ from \emph{cvx-fin-notes.tex}.

\item[$6 \implies 5$] Trivial
\item[$5 \implies 1$] Let $\trim(\Gamma)$ be finite.
Then there is a nondegenerate, nonredundant, finite property $\gamma$ so that $\{ \gamma_r : r \in \R' \} = \strip(\gamma) = \trim(\Gamma)$.

Then we want to show $\Gamma$ embeds $\gamma$.
Since we know $\card(\trim(\Gamma)) = \card(\R')$ is finite and $\trim(\Gamma)$ is a subset of the codomain of $\Gamma$, $\R$, then $\card(\R')) \leq \card(\R)$ \jessie{$?= \card(\reals^d)$... again assuming $\R$ is a \emph{convex} subset of $\reals^d$}, there is an injection $\varphi : \R' \to \R$ from Folland Section 0.

Consider some $r \in \gamma(p)$.
Since there is an injection $\varphi$ from $\R' \to \R$, then $r \in \gamma(p) \implies \varphi(r) \in \varphi(\gamma(p)) = \Gamma(p)$.
\jessie{Hand wavy...}

@@

Therefore, $5 \implies 1$.

@@

\begin{lemma}\label{lem:pos-measure-iff-nonempty-int}
Let $S$ be a convex set.
$S$ has nonempty interior $\implies$ has positive Lebesgue measure.
\end{lemma}
\begin{proof}

%%%\begin{enumerate}
%%%	\item[$\implies$]
	Let $S$ have nonempty interior.
	We know that as $\inter{S}$ and $bd(S)$ are disjoint, $\mu(S) = \mu(\inter{S}) + \mu(bd(S))$.
	Since the interior of $S$ is nonempty, there is an open ball $B \subseteq S$ such that $x \in B \implies x \in \inter S$.
	As Lebesgue measure in $\reals^n$ is a product measure and every open interval has positive measure, then the Lebesgue measure of the ball $B$ is the product of finitely many positive measures, and therefore positive.
	As $0 < \mu(B) \leq \mu(\inter S) \leq \mu(S)$ by monotonicity, we see that $S$ with nonempty interior has positive measure.

%%%	\item[$\impliedby$]  		
%%%	Follows from the contrapositive and definition of measure.
%%%%		That is, $\mu(S) = \mu(\inter{S}) + \mu(bd(S)) = \mu(\inter{S})$.\footnote{Since $S$ is convex, $\mu(bd(S)) = 0$ according to \url{https://math.stackexchange.com/questions/613408/when-does-the-boundary-have-measure-zero}.}
%%%%		As $\inter{S} = \emptyset \implies \mu(\inter{S}) = 0$, and hence does not have positive measure.

%%%		\jessie{Use Lemma~\ref{lem:nonempty-inter-iff-aff-ind}.}
%%%	\end{enumerate}
\end{proof}

\begin{lemma}\label{lem:nonempty-inter-iff-aff-ind}
	Let $\theta$ be a [closed] convex level set.
	$\theta$ is full dimensional $\iff$ there are distributions $p_1, p_2, \ldots, p_n \in \theta$ so that $p_1, p_2, \ldots, p_n$ are affinely independent.
\end{lemma}
\begin{proof}
	Recall that $n := |\Delta(\Y)|$, so we say the simplex is $n$-dimensional.
	\begin{enumerate}
		\item [$\implies$]
		Assume the level set $\theta$ is full dimensional.
		Then there exists a $p \in \inter{\theta}$ and for every $q \in B(\epsilon, p)$, we know that $q \in \theta$.
		Take $q_i = p + \epsilon e_i$\footnote{$e_i$ is the vector with $1$ in the $i^{th}$ component, and is $0$ elsewhere} in each of the $n$ dimensions, and since $\theta$ is full dimensional and closed, we know each $q_i \in\theta$.
		As $\{e_i\}_{i=1}^n$ is affinely independent (it forms the standard basis for $\reals^n$), so then is $\{q_i\}_{i=1}^n$ as we are invariant to additive and multiplicative shifts.

@@

\begin{conjecture}\label{conj:prec-if-embeds}
	If the elicitable property $\Gamma$ embeds the nondegenerate \jessie{elicitable?} finite $\gamma$, then $\gamma \preceq \Gamma$ and $\trim(\Gamma) = \strip(\gamma)$.
\end{conjecture}
\begin{proof}

@@

	To show $\gamma \preceq \Gamma$, we know that since $\Gamma$ embeds $\gamma$, there is an injection $\varphi : \R' \to \R$ so that $r' \in \gamma(p) \iff \varphi(r') \in \Gamma(p)$.
	Therefore, we know for all $r' \in \R'$ that $\varphi(r') \in \R$, therefore, $\varphi(\R') \subseteq \R$.
	As $\varphi(\R') \subseteq \R$ and $\gamma$ is nondegenerate, we know that for all $r \in \R'$, we know $\gamma \preceq \Gamma$.
	Therefore, $\gamma(p) = \Gamma(p) \cap \R'$

	Finally, we show $\gamma$ is nonredundant. 
	For contradiction, suppose there are two reports $r, r' \in \R'$ so that $r \neq r'$ and $\gamma_r \subseteq \gamma_{r'}$.
	Since $\Gamma$ embeds $\gamma$, for any $p \in \P$, we know $p \in \gamma_r \iff p \in \Gamma_{\varphi(r)}$.
	Therefore, $\Gamma_{\varphi(r)} \subseteq \Gamma_{\varphi(r')}$, where $\varphi(r) \neq \varphi(r')$ as $\varphi$ is an injection.

	Since $\gamma_r \subseteq \gamma_{r'}$, then $\gamma_r \cap \gamma_{r'} = \gamma_r$.
	We know $p \in \gamma_r$ and $\gamma_{r'} \iff p \in \Gamma_{\varphi(r)}$ and $\Gamma_{\varphi(r')} \iff p \in \Gamma_{\lambda \varphi(r) + (1-\lambda) \varphi(r')}$ for all $\lambda \in [0,1]$ by elicitability of $\Gamma$ and closure of subgradient sets of $G_{\Gamma}$.
	Since the reals are (uncountably) infinite, we then contradict the finiteness of $\gamma$.
	Therefore, we conclude $\gamma$ must be nonredundant.
	\jessie{I don't think this is correct... I think we need convexity of $\R$ in order to claim this, which we technically don't assume.}
